Implementation of Full Synchronous Composition of Discrete Event Models Using IEC 61499 Function Blocks

### 2005-01-31 doc/Sync_Execution_with_IEC61499/paper.tex
*** empty log message ***

\author{\authorblockN{Goran Cengic, Knut Åkesson and
    Bengt Lennartson}
\authorblockA{
  Automation Laboratory\\
  Department of Signals and Systems\\
  Chalmers University of Technology\\
  Sweden
}
\and
\authorblockN{Chengyin Yuan and Placid Ferreira}
\authorblockA{
  Department of Mechanical and Industrial Engineering\\
  University of Illinois at Urbana-Champaign\\
  USA
}
}
\maketitle
\begin{abstract}
  A new method for implementation of synchronous composition
  is presented. Using an emerging standard, IEC 61499, the
  method enables efficient development of distributed
  control applications. It also brings together supervisory
  control theory with distributed control application
  development since synchronous composition of discrete
  event models is used to model interaction between systems
  in supervisory control theory. The method is demonstrated
  using a simple example before the formulation for the
  general case is presented. A new structure of control
  applications suitable for application of supervisory
  control theory to control application development is
  presented also.
\end{abstract}

@@

There are existing methods for implementation of
synchronization operator using PLC programming techniques
\cite{hfl:syn:1999,hfl:mod:2001,fh:plc:1998} but the
distribution is hard to achieve in that programming
environment. A new method presented in this paper uses the
emerging standard for distributed control systems to remedy
that.

@@

In the supervisory control theory, the plant is assumed to
generate all events and the purpose of the supervisor is to
prevent the plant from generating events that will violate
the specification. The assumption that the plant generates
all events might seem like a restriction. However, it is
often possible to introduce a controller which will initiate
the events between the supervisor and the resource. Treating
the controller as part of the plant will make the assumption
that all events are generated by the plant a natural one. In
Fig.~\ref{fig:sup_plant} such configuration is shown.
\begin{figure}
  \centering \includegraphics[scale=0.6]{figures/sup_plant}
  \caption{Supervisory control configuration.}
  \label{fig:sup_plant}
\end{figure}

The supervisor, S in Fig.~\ref{fig:sup_plant}, has to
prevent the plant from breaking the specification and from
getting blocked in a state from where it can not reach a
marked state by using as few restrictions on the plant as
possible. By observing the observable events of the plant
and disabling the controllable events of the plant, the
supervisor decides when, and if, it has to limit the
activity of the plant in order to keep the plant's behavior
within the specification. An important result of the
supervisory control theory is that such a supervisor can be
calculated from the models of the plant and specification so
that
$$S||P'$$
gives the system behavior.

@@

Consider a system containing two plants, \texttt{P1} and
\texttt{P2}, shown in Fig.~\ref{fig:plants}.These plant
models may represent two manufacturing resources with a
1-place buffer in-between. All events in the models are
controllable and observable. The desired behavior of the
system is that event \texttt{b} must always precede event
\texttt{c}, which means that a product must leave buffer
before another one can enter. The supervisor for this
specification can be generated and is shown in
Fig.~\ref{fig:sup}. In cases where the specification is
more complex it needs to be modeled using the automaton and
provided for the supervisor generation.
\begin{figure}[b]
  \centering 
  \fig[scale=0.55]{\texttt{P1}}{figures/P1_aut} 
  \fig[scale=0.55]{\texttt{P2}}{figures/P2_aut}
  \caption{Discrete event models of the plants.}
  \label{fig:plants}
\end{figure}
\begin{figure}[b]
  \centering 
  \fig[scale=0.55]{\texttt{S}}{figures/Sp_aut}
  \caption{The model of the supervisor.}
  \label{fig:sup}
\end{figure}
\begin{figure}[t]
  \centering 
  \fig[scale=0.45]{\texttt{SYS}}{figures/Sys_aut}
  \caption{The model of possible system behavior while satisfying
    the desired system behavior, $S||P_1||P_2$.}
  \label{fig:sys}
\end{figure}
Full synchronous execution of these models gives the system
that behaves as freely as possible while fulfilling the
specification. The resulting model of the system is shown in
Fig.~\ref{fig:sys}.

% IEC 61499 intro
The emerging standard IEC 61499 specifies a generic
architecture for measurement and control of industrial
processes. The architecture is based on functional software
units called function blocks. The most common type of
function block in an application is basic function block and
it is shown in Fig.~\ref{fig:61499_intro} using the
graphical representation. This type of block executes
algorithms based on the arriving events and generates new
events that are passed on when the algorithms finish
execution. The algorithms use data associated with incoming
events to update the internal variables and produce the
output data associated with the outgoing events. Which
algorithms are invoked is controlled by the execution
control chart (ECC) which is an automaton with algorithms
associated with different states. When a state is entered
the algorithms associated with it are executed once and the
ECC stays in the state until a condition for leaving is
fulfilled. The conditions upon which transitions occur are
boolean expressions involving event input variables,
internal variables, input variables, and output variables.
When the expression evaluates to \texttt{true} the
transition occur. Special case is transition labeled with
\texttt{1}, it is taken as soon as the algorithms finish
execution.
\begin{figure}
  \centering
  \includegraphics[scale=0.4]{figures/basic_block}
  \caption{Anatomy of a basic function block type.}
  \label{fig:61499_intro}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{figures/P1}\\(a)
  \vskip 5mm
  \includegraphics[scale=0.5]{figures/P2}\\(b)
  \vskip 5mm
  \includegraphics[scale=0.5]{figures/Sp}\\(c)
  \caption{The function block types implementing (a)
    plant 1, (b) plant 2, and (c) supervisor.}
  \label{fig:systemFBs}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[scale=0.5]{figures/SYNC}
  \caption{The function block type for the synchronization.}
  \label{fig:syncFB}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[scale=0.6]{figures/sync_ecc}
  \caption{The execution control chart for the
    synchronization function block type.}
  \label{fig:ecc_sync}
\end{figure}

@@

The control application connecting all of the function
blocks together is shown in Fig.~\ref{fig:app}. Small
circles in the figure are event multiplexer and
demultiplexer. Event multiplexer generates an event on the
output every time an event arrives at any one of the inputs.
Event demultiplexer does the opposite and generates an event
on all of the outputs whenever an event arrives at the
input.
\begin{figure}
  \centering
  \includegraphics[scale=0.45]{figures/app}
  \caption{The control application for the example.}
  \label{fig:app}
\end{figure}

@@

Today implementation of a distributed control application is
done using application structure similar to the structure in
Fig.~\ref{fig:old_app_struct}. Supervisor \texttt{S} in
the application is manually crafted using informal
specifications of the system behavior, and transfered to the
PLC-code. For each application the communication between the
supervisor and the plants has to be implemented in a
application specific way since the existing standards for
communication between PLCs are at a low level of
abstraction. Using IEC 61499 the communication part of the
application is abstracted far away from the hardware level
and the control engineer is only considering the event and
data passing between the functional parts of the
application, i.e. function block instances. When the system
plants are upgraded or changed the procedure of supervisor
crafting has to be done once again and very small parts of
the code from the application already implemented are
reused, i.e. the communication parts.
\begin{figure}
  \centering 
  \includegraphics[scale=0.6]{figures/old_app_struct}
  \caption{The structure of distributed control applications today.}
  \label{fig:old_app_struct}
\end{figure}

>>>>

\author{\authorblockN{Goran Cengic, Knut Åkesson and
    Bengt Lennartson}
\authorblockA{
  Automation Laboratory\\
  Department of Signals and Systems\\
  Chalmers University of Technology\\
  Sweden
}
\and
\authorblockN{Chengyin Yuan and Placid Ferreira}
\authorblockA{
  Department of Mechanical and Industrial Engineering\\
  University of Illinois at Urbana-Champaign\\
  USA
}
}
\maketitle
\begin{abstract}
  One reason that the development of distributed control
  system is hard is that the communication primitives are at
  a low abstraction level. To put less burden on the
  developer it would be an advantage to have support for
  high-level communication primitives. One such primitive is
  the synchronous composition operator. This operator can be
  used to model the interaction between an arbitrary number
  of concurrently executing applications. This paper shows
  how to implement the synchronous composition operator
  using an emerging communication standard for distributed
  control systems, IEC 61499. The operator is used in the
  supervisory control theory, a framework for automatic
  generation of provable correct supervisors. In
  manufacturing systems, for example, the supervisor might
  have to interact with a physically distributed plant. This
  paper concludes therefore by showing how to use the
  proposed implementation of the synchronous composition
  operator to execute a supervisor against a physically
  distributed plant.

\end{abstract}

% supervisory control theory, discrete event systems,
% manufacturing systems, synchronous composition, IEC 61499,
% Distributed Control Systems

@@

There are existing methods for implementation of
synchronization operator using PLC programming techniques
\cite{hfl:syn:1999,hfl:mod:2001,fh:plc:1998,a:met:2002} but
the distribution is hard to achieve in that programming
environment. A new method presented in this paper uses the
emerging standard for distributed control systems to remedy
that.

@@

In the supervisory control theory, the plant is assumed to
generate all events and the purpose of the supervisor is to
prevent the plant from generating events that will violate
the specification. The assumption that the plant generates
all events might seem like a restriction. However, it is
often possible to introduce a controller which will initiate
the events between the supervisor and the resource. Treating
the controller as part of the plant will make the assumption
that all events are generated by the plant a natural one. In
Fig.~\ref{fig:sup_plant} such configuration is shown.
\begin{figure}
  \centering 
  \includegraphics[scale=0.6]{figures/sup_plant}
  \caption{Supervisory control configuration with a resource
  \texttt{R}, controller \texttt{C}, and a supervisor \texttt{S}.}
  \label{fig:sup_plant}
\end{figure}

The supervisor, \texttt{S} (see Fig.~\ref{fig:sup_plant}),
has to prevent the plant, \texttt{P}, from breaking the
specification and from getting blocked in a state from where
it can not reach a marked state by using as few restrictions
on the plant as possible. By observing the observable events
of the plant and disabling the controllable events of the
plant, the supervisor decides when, and if, it has to limit
the activity of the plant in order to keep the plant's
behavior within the specification. An important result of
the supervisory control theory is that such a supervisor can
be calculated from the models of the plant and specification
so that
$$S||P$$
gives the system behavior.

@@

Consider a system containing two plants, \texttt{P1} and
\texttt{P2}, shown in Fig.~\ref{fig:plants}.These plant
models may represent two manufacturing resources with a
1-place buffer in-between. All events in the models are
controllable and observable. The desired behavior of the
system is that event \texttt{b} must always precede event
\texttt{c}, which means that a product must leave buffer
before another one can enter. The supervisor for this
specification can be generated and is shown in
Fig.~\ref{fig:sup}. In cases where the specification is
more complex it needs to be modeled using the automaton and
provided for the supervisor generation.
\begin{figure}[b]
  \centering 
  \fig[scale=0.55]{\texttt{P1}}{figures/P1_aut} 
  \fig[scale=0.55]{\texttt{P2}}{figures/P2_aut}
  \caption{Discrete event models of the plants. These may
    represent two manufacturing resources with a 1-place
    buffer in-between.}
  \label{fig:plants}
\end{figure}
\begin{figure}[b]
  \centering 
  \fig[scale=0.55]{\texttt{S}}{figures/Sp_aut}
  \caption{The model of the supervisor that make shure that
    event \texttt{b} always occurs before event \texttt{c}.}
  \label{fig:sup}
\end{figure}
\begin{figure}[t]
  \centering 
  \fig[scale=0.45]{\texttt{SYS}}{figures/Sys_aut}
  \caption{The model of possible system behavior while satisfying
    the desired system behavior, $S||P_1||P_2$.}
  \label{fig:sys}
\end{figure}
Full synchronous execution of these models gives the system
that behaves as freely as possible while fulfilling the
specification. The resulting model of the system is shown in
Fig.~\ref{fig:sys}.

% IEC 61499 intro
The emerging standard IEC 61499 specifies a generic
architecture for measurement and control of industrial
processes. The architecture is based on functional software
units called function blocks. The most common type of
function block in an application is basic function block and
it is shown in Fig.~\ref{fig:61499_intro} using the
graphical representation. This type of block executes
algorithms based on the arriving events and generates new
events that are passed on when the algorithms finish
execution. The algorithms use data associated with incoming
events to update the internal variables and produce the
output data associated with the outgoing events. Which
algorithms are invoked is controlled by the execution
control chart (ECC) which is an automaton with algorithms
associated with different states. When a state is entered
the algorithms associated with it are executed once and the
ECC stays in the state until a condition for leaving is
fulfilled. The conditions upon which transitions occur are
boolean expressions involving event input variables,
internal variables, input variables, and output variables.
When the expression evaluates to \texttt{true} the
transition occur. Special case is transition labeled with
\texttt{1}, it is taken as soon as the algorithms finish
execution.
\begin{figure}
  \centering
  \includegraphics[scale=0.65]{figures/basic_block}
  \caption{Anatomy of a basic function block type.}
  \label{fig:61499_intro}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.45]{figures/P1}\\(a)
  \vskip 5mm
  \includegraphics[scale=0.45]{figures/P2}\\(b)
  \vskip 5mm
  \includegraphics[scale=0.45]{figures/Sp}\\(c)
  \caption{The function block types implementing (a)
    plant 1, (b) plant 2, and (c) supervisor.}
  \label{fig:systemFBs}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[scale=0.5]{figures/SYNC}
  \caption{The function block type for synchronization.}
  \label{fig:syncFB}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[scale=0.45]{figures/sync_ecc}
  \caption{The execution control chart for the
    synchronization function block type.}
  \label{fig:ecc_sync}
\end{figure}

@@

The control application connecting all of the function
blocks together is shown in Fig.~\ref{fig:app}. Small
circles in the figure are event multiplexer and
demultiplexer. Event multiplexer generates an event on the
output every time an event arrives at any one of the inputs.
Event demultiplexer does the opposite and generates an event
on all of the outputs whenever an event arrives at the
input.
\begin{figure}
  \centering
  \includegraphics[scale=0.45]{figures/app}
  \caption{The control application for the example with
  function block instances for plants, \texttt{P1} and
  \texttt{P2}, supervisor, \texttt{S}, and synchronization
  block, \texttt{SYNC}.}
  \label{fig:app}
\end{figure}

@@

Today most implementations of distributed control
applications are done using application structure similar to
the structure in Fig.~\ref{fig:old_app_struct}. The
controller, \texttt{C}, in application is manually crafted
using informal specifications of the system behavior, and
transfered to the PLC-code. For each application the
communication between the supervisor and the plants has to
be implemented in an application and vendor specific way
since the existing standards for communication between PLCs
are at a low level of abstraction. Using IEC 61499 the
communication part of the application is abstracted away
from the hardware level and the control engineer is only
considering the event and data passing between the
functional parts of the application, i.e. function block
instances. When the system plants are upgraded or changed
the procedure of supervisor crafting has to be done once
again and very small parts of the code from the application
already implemented are reused, i.e. the communication
parts.
\begin{figure}
  \centering 
  \includegraphics[scale=0.6]{figures/old_app_struct}
  \caption{The structure of distributed control applications
  today with plants, \texttt{P1} and \texttt{P2}, and
  controller, \texttt{C}.}
  \label{fig:old_app_struct}
\end{figure}
